Multiple testing with logical dependencies

### 2013-08-15 paper/draft.tex
Added proof for hierarchical result

\begin{theorem}
  Consider a family of null hypotheses for which conditions
  (\ref{cond:matroida}-\ref{cond:matroidb}) hold. Then an index set $I \subseteq S$ is
  weakly exhaustive if and only if $I$ is a flat in $M$.
\end{theorem}

@@

\begin{theorem}
  Consider a hierarchical testing procedure that tests $H_{i_r}$ at
  level $\alpha$ according to a prefixed order $(i_1,...,i_n)$. Assume
  that hypotheses with indeces $i_1$ through $i_k$ have been
  rejected. If for $k+1$ to $k+l$ there is $i_s$, $s \leq k$ such that
  \begin{equation}
    \label{eq:condh}
    \rho \{i_s,i_{k+1}\} = \rho \{i_s,i_{k+1}...,i_{k+l}\},
  \end{equation}
  hypotheses each $H_{i_{k+1}}$ through   $H_{i_{k+l}}$ can be tested
  simultaneously at full level $\alpha$. 
\end{theorem}

\begin{proof}
  Due to \ref{eq:condh} for any $k+1 \leq m' < m \leq k+l$, $i_m \in
  \sigma \{i_s,i_{m'}\}$ and consequently also $i_s \in \sigma
  \{i_m,i_{m'}\}$. Therefore $i_s \in \sigma J$ for all $J$ with
  $i_m,i_{m'} \in J$. Since $p_s \leq \alpha$ all intersection
  hypotheses corresponding to exhaustive sets containing $i_{m}$
  and $i_{m'}$ for all $k+1 \leq m' < m$ have been rejected. Such that
  $p_{i_{m}} \leq \alpha$ rejects $H_{i_{m}}$.
\end{proof}

>>>>

\begin{theorem}
  \label{the:strictly.exhaustive}
  Consider a family of null hypotheses for which conditions
  (\ref{cond:matroida}-\ref{cond:matroidb}) hold. Then an index set $I \subseteq S$ is
  weakly exhaustive if and only if $I$ is a flat in $M$.
\end{theorem}

@@

% \begin{theorem}
%   Consider a hierarchical testing procedure that tests $H_{i_r}$ at
%   level $\alpha$ according to a prefixed order $(i_1,...,i_n)$. Assume
%   that hypotheses with indeces $i_1$ through $i_k$ have been
%   rejected. If for $k+1$ to $k+l$ there is $i_s$, $s \leq k$ such that
%   \begin{equation}
%     \label{eq:condh}
%     \rho \{i_s,i_{k+1}\} = \rho \{i_s,i_{k+1}...,i_{k+l}\},
%   \end{equation}
%   hypotheses each $H_{i_{k+1}}$ through   $H_{i_{k+l}}$ can be tested
%   simultaneously at full level $\alpha$. 
% \end{theorem}

% \begin{proof}
%   Due to \ref{eq:condh} for any $k+1 \leq m' < m \leq k+l$, $i_m \in
%   \sigma \{i_s,i_{m'}\}$ and consequently also $i_s \in \sigma
%   \{i_m,i_{m'}\}$. Therefore $i_s \in \sigma J$ for all $J$ with
%   $i_m,i_{m'} \in J$. Since $p_s \leq \alpha$ all intersection
%   hypotheses corresponding to exhaustive sets containing $i_{m}$
%   and $i_{m'}$ for all $k+1 \leq m' < m$ have been rejected. Such that
%   $p_{i_{m}} \leq \alpha$ rejects $H_{i_{m}}$.
% \end{proof}

\begin{theorem}
  Consider a system of logically dependent hypotheses $\mathscr{H}$
  with dependence relation $\sim$ and associated matroid
  $M(\mathscr{H})$. Assume that $M(\mathscr{H})$ neither contains
  loops nor parallels. Consider a hierarchical testing procedure that
  tests $H_{i}$ at level $\alpha$ according to a prefixed
  order. Without loss of generality assume that $H_i$ are indexed
  according to the fixed rejection sequence $(1,...,n)$. Given 
  that hypotheses with indeces $1$ through $r$ have been
  rejected and if for $r+1$ to $r+s$ holds
  \begin{equation}
    \label{eq:cond.hierarchical}
    \rho \{1,...,r\} + 1 = \rho \{1,...,r+1\} = \rho \{1,...,r+s\},
  \end{equation}
  hypotheses $H_{r+1}$ through $H_{r+s}$ can be tested simultaneously
  at full level $\alpha$.
\end{theorem}

\begin{proof}
  We show by that given $p_{r+k} < \alpha$ all $E \in \mathscr{L}_{r+k} =
  \{I \in \mathscr{L}: r+k \in I\}$ are rejected by the stepwise
  procedure. First, notice that for all $E$ such that $\min E = r+k$
  rejection follows immediately and if $\min E \leq r$ rejection
  follows by rejection of $H_i$ for $i \leq r$. It remains to be shown
  that the stepwise procedure also rejects $H_E$ for $r < \min E <
  r+k$.

  First we show by contradiction that some $k \leq t$ we have $i \in
  \sigma \{r+l,r+k\}$ for some $i \leq r$ and all $1 \leq l < k$.  We
  need to show that for all $s,t$, $k+1 < s < t \leq k+l$, follows
  $i_s \sim K \cup i_t$:

  Assume that there is an $l$ such that there is on $i \leq r$ with $i
  \in \sigma \{r+l,r+k\}$. Consequently for all $i \leq r$ $i \nsim
  \{r+l,r+k\}$. Since $M(\mathscr{H})$ has no loops or parallels also
  $r+l \nsim r+k$. Consequently, for any maximally independent set in
  $I \subset \{1,...,r\}$, $I \cup \{r+l, r+k\}$ is independent and
  has cardinality $\rho\{1,...,r\}+2$ which contradicts
  (\ref{eq:cond.hierarchical}).

  Above argument can be repeated for any set $J \cup r+k$ where $J
  \subseteq \{r+1,...,r+k-1\}$. Therefore, we have $i \in E$, $i \leq
  r$ for all exhaustive set containing $r+k$ with $r < \min E <
  r+k$. Rejection of these sets follows from rejection of $H_i$, $i
  \leq r$, which completes the proof. 
\end{proof}

Note that the condition that $M(\mathscr{H})$ contains neither loops
nor parallels is not a serious restriction. A loop in terms of systems
of hypotheses would be a set in the parameter space that contains all
hypotheses. Thich corresponds to the alternative that at least one
null hypotheses is true which is the domain of intersection-union
tests. The condition that $M(\mathscr{H})$ does not contain parallels
is more restrictive, however, considering that the absence of
parallels is a sufficient condition such that the correspondence
between flats of the matroid and strictly exhaustive sets (Theorem
\ref{the:strictly.exhaustive}) this is also not serious. Once we turn
to one-sided problems where we will use oriented matroids to describe
the structure of systems of null hypotheses these problems will seem
even less problematic.
